On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-11-28 manuscripts/PNAS/second review/resub/chazthm_pnas_resub.tex
cjh edits from muddy waters

\begin{definition}\label{maindef}
Fix $Y = \{ \mathbf{y}_1, \ldots, \mathbf{y}_N\} \subset \mathbb{R}^n$. We say $Y$ has a \textbf{$k$-sparse representation in $\mathbb{R}^m$} if there exists a matrix $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ such that $\mathbf{y}_i = \mathbf{A}\mathbf{x}_i$ for all $i$. 
This representation is \textbf{stable} if for every $\delta_1, \delta_2 \geq 0$, there exists some $\varepsilon = \varepsilon(\delta_1, \delta_2)$ that is strictly positive for positive $\delta_1$ and $\delta_2$ such that if $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^m$ satisfy:
\begin{align*}
	\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \varepsilon(\delta_1, \delta_2),\ \ \   \text{for all $i$},
\end{align*}
then there is some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that for all $i, j$:
\begin{align}\label{def1}
\|\mathbf{A}_j - \mathbf{BPD}_j\|_2 \leq \delta_1 \ \ \text{and} \ \ \|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\overline x}_i\|_1 \leq \delta_2.
\end{align}
\end{definition}

To see how Def. \ref{maindef} directly relates to Prob. \ref{InverseProblem}, suppose that $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ and fix $\delta_1, \delta_2$ to be the desired recovery accuracy in \eqref{def1}. Consider now any dataset $Z$ generated as in \eqref{LinearModel} with $\eta \leq \frac{1}{2} \varepsilon(\delta_1, \delta_2)$. Then from the triangle inequality, it follows that any matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^m$ solving Prob.~\ref{InverseProblem} are necessarily within $\delta_1, \delta_2$ of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$, respectively.

@@

\begin{problem}\label{OptimizationProblem}
Find $\mathbf{B}$ and \mbox{$\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^{\overline m}$} that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\overline x}_i\|_0 \ \ 
\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta, \ \text{for all $i$}.
\end{align}
\end{problem}
We then sketch the proof of a probabilistic extension of Thm.~\ref{DeterministicUniquenessTheorem} to random data and dictionaries (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}). Finally, in Sec.~\ref{Discussion}, we discuss both theoretical and practical applications of our main mathematical findings.

@@

Next, we describe a quantitative version of the spark condition. The \emph{lower bound} of a matrix $\mathbf{M} \in \mathbb R^{n \times m}$ is the largest $\alpha$ such that \mbox{$\|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$ \cite{Grcar10}. By compactness of the unit sphere, every injective linear map has nonzero lower bound; hence, if $\mathbf{M}$ satisfies \eqref{SparkCondition}, then each submatrix formed from $2k$ of its columns or less has a strictly positive lower bound.

@@

Every $n \times \overline m$ matrix $\mathbf{B}$ for which there are $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N$ satisfying \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \varepsilon$} for all $i$ has $\overline m \geq m$ and, provided $\overline m < mr/(r-1)$,
\begin{align}\label{Cstable}
\|\mathbf{A}_j- \mathbf{B}_{\overline J} \mathbf{PD}_j\|_2 \leq C_1 \varepsilon, \ \ \text{for all } j \in J,
\end{align}
%
for some nonempty $\overline J \subseteq [\overline m]$ and $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$, permutation matrix $\mathbf{P}$, and and invertible diagonal matrix $\mathbf{D}$.

@@

One can also easily verify that for every $k < m$, there are regular $k$-uniform hypergraphs with the SIP besides the obvious $\mathcal{H} = {[m] \choose k}$; for instance, take $\mathcal{H}$ to be the consecutive intervals of length $k$ in some cyclic order on $[m]$. In this case, a direct consequence of Cor.~\ref{DeterministicUniquenessCorollary} is rigorous verification of the lower bound \mbox{$N = m(k-1){m \choose k} + m$} for sufficient sample size from the introduction. In many cases, however, the SIP is achievable with far fewer supports. For example, when $k = \sqrt{m}$, simply take $\mathcal{H}$ to be the $2k$ rows and columns formed by arranging $[m]$ in a square grid.

@@

We furthermore note that unlike in previous works \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}, the matrix $\mathbf{A}$ need not satisfy the spark condition to be recoverable from data. As a simple example for $k=2$, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ where $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$, and take $\mathcal{H}$ to be all consecutive pairs of $[m]$ arranged in cyclic order. Then, dictionary $\mathbf{A}$ satisfies the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} underlying \eqref{Cstable} without satisfying \eqref{SparkCondition}. %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
%This weakening allows for a practical (polynomial) amount of data to still guarantee a stable dictionary in the case where the set of sparse supports for the $\mathbf{\overline x}_i$ is known to have a size that grows polynomially in $m$ and $k$.

@@

\begin{proposition}\label{neuroprop}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse linear coding (e.g., \cite{rehnsommer2007, rozell2007neurally}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}

@@

\begin{definition}\label{FriedrichsDefinition}
For a collection of real subspaces $\mathcal{V} = \{V_i\}_{i=1}^\ell$, define $\xi^2 := 1$ when $|\mathcal{V}| = 1$, and otherwise:
\begin{align}\label{xi}
\xi^2(\mathcal{V}) := 1 -  \max \prod_{i=1}^{\ell-1} \sin^2  \theta \left(V_i, \cap_{j>i} V_j \right) ,
\end{align} 
%
where the maximum is taken over all orderings 
%\footnote{We modify the quantity in \cite{Deutsch12} in this way since the subspace ordering is irrelevant to our purpose.} 
of the $V_i$ and the angle $\theta \in (0,\frac{\pi}{2}]$ is defined implicitly as \cite[Def.~9.4]{Deutsch12}:
\begin{align*}
\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \substack{ \mathbf{u} \in U \cap (U \cap W)^\perp, \ \|\mathbf{u}\|_2 \leq 1 \\ \mathbf{w} \in W \cap (U \cap W)^\perp, \  \|\mathbf{w}\|_2 \leq 1 } \right\}.
\end{align*}
\end{definition}
Note that $\theta \in (0,\frac{\pi}{2}]$ implies $0 < \xi \leq 1$, and that $\xi(\mathcal{V}') \leq \xi(\mathcal{V})$ when $\mathcal{V}' \subseteq \mathcal{V}$.\footnote{We acknowledge the counter-intuitive property that $\theta =  \pi/2$ when $U \subseteq W$.}  
The constant $C_2$ in Lem.~\ref{MainLemma} is then:  
\begin{align}\label{Cdef2}
	C_2(\mathbf{A}, \mathcal{H}) := \frac{ (r+1) \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ 1- \max_{\mathcal{G} \in {\mathcal{H} \choose r+1}} \xi( \bm{\mathcal{A}}_\mathcal{G} ) },
\end{align}
% C_2(\mathbf{A}, \mathcal{H}) := \frac{ (r+1) \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ \min_{\mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1}} \xi( \{ \bm{\mathcal{A}}_S \}
which we remark is consistent with the assumption on $C_1$ in the proof of the case $k=1$ at the beginning of this section.

@@

Since there are more than $(k-1){\overline m \choose k}$ vectors $\mathbf{x}_i$ supported in $S$, the pigeonhole principle gives $\overline S \in {[\overline m] \choose k}$ and a set of $k$ indices $K \subseteq I(S)$ with all $\mathbf{\overline x}_i$, $i \in K$, supported in $\overline S$.

@@

\begin{lemma}\label{spanIntersectionLemma}
If $f: U \to V$ is an injective function, then $f\left(\cap_{i=1}^\ell U_i \right) =  \cap_{i=1}^\ell f\left(U_i\right)$ for any $U_1, \ldots, U_\ell \subseteq U$.
%In particular, if for some $E \in 2^{[m]}$ the map $M \in \mathbb{R}^{n \times m}$ is injective on $\cup_{S \in \mathcal{H}} \text{span}\{e_i\}_{i \in S}$ then $\text{span}\{ M_{\cap_{S \in \mathcal{H}} S} \} = \cap_{S \in \mathcal{H}} \text{span}\{M_S\}$.
\end{lemma}
\begin{proof}
By induction, it is enough to prove the trivial case $\ell = 2$. %Clearly, for any map $f$, if $w \in f(U \cap V)$ then $w \in f(U)$ and $w \in f(V)$, hence $w \in f(U) \cap f(V)$. If $w \in f(U) \cap f(V)$ then $w \in f(U)$ and $w \in f(V)$, hence $w = f(u) = f(v)$ for some $u \in U$ and $v \in V$, implying $u = v$ by injectivity of $f$. Hence $u \in U \cap V$, and $w \in f(U \cap V)$.
\end{proof}
% \vspace{-.4 cm}

\begin{lemma}\label{DistanceToIntersectionLemma}
Let $\mathcal{V} = \{V_i\}_{i=1}^k$ be a set of two or more subspaces of $\mathbb{R}^m$ and let $V = \cap_{i = 1}^k V_i$. For  $\mathbf{x} \in \mathbb{R}^m$, we have (recall Defs.~\ref{dDef}~\&~\ref{FriedrichsDefinition}):
\begin{align}\label{DTILeq}
\text{\rm dist}(\mathbf{x}, V) \leq \frac{1}{1 - \xi(\mathcal{V})} \sum_{i=1}^k \text{\rm dist}(\mathbf{x}, V_i).
\end{align}
\end{lemma}
\vspace{-.3 cm}
\begin{proof} 
Recall the orthogonal projection onto the subspace $V \subseteq \mathbb{R}^m$ is the mapping $\Pi_V: \mathbb{R}^m \to V$ that associates with each $\mathbf{x}$ its unique nearest point in $V$; i.e., $\|\mathbf{x} - \Pi_V\mathbf{x}\|_2 = \text{\rm dist}(\mathbf{x}, V)$.
Next, observe:
\begin{align}\label{f}
\|\mathbf{x} - \Pi_V\mathbf{x}\|_2 &\leq \|\mathbf{x} - \Pi_{V_k} \mathbf{x}\|_2 + \|\Pi_{V_k}  \mathbf{x} - \Pi_{V_k}\Pi_{V_{k-1}}\mathbf{x}\|_2 \nonumber \\
&\ \ \ + \cdots + \|\Pi_{V_k} \Pi_{V_{k-1}}\cdots \Pi_{V_1} \mathbf{x} - \Pi_V \mathbf{x}\|_2 \nonumber \\
&\leq \|\Pi_{V_k}\cdots\Pi_{V_{1}} \mathbf{x} - \Pi_V \mathbf{x}\|_2 + \sum_{\ell=1}^k \|\mathbf{x} - \Pi_{V_{\ell}} \mathbf{x}\|_2,
\end{align}
using the triangle inequality and that the spectral norm satisfies $\|\Pi_{V_{\ell}}\|_2 \leq 1$ for all $\ell$ (since $\Pi_{V_{\ell}}$ are orthogonal projections).

@@

Finally, substituting this into \eqref{f} and rearranging produces \eqref{DTILeq} after substituting $\xi(\mathcal{V})$ for $z$.
\end{proof}
\begin{lemma}\label{NonEmptyLemma} 
Fix an $r$-regular hypergraph $\mathcal{H} \subseteq 2^{[m]}$ satisfying the SIP. If the map $\pi: \mathcal{H} \to 2^{[\overline m]}$ has $\sum_{S \in \mathcal{H}} |\pi(S)| \geq \sum_{S \in \mathcal{H}} |S|$ and:
\begin{align}\label{cond}
	|\cap \pi(\mathcal{G})| \leq |\cap \mathcal{G} |,\ \ \   \text{for } \mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1},
\end{align}
%
then $\overline m \geq m$, and if $\overline m < mr / (r-1)$ then the map $i \mapsto \cap \pi(\sigma(i))$ is an injective map to $[\overline m]$ from some $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$. %In particular, if $\overline m = m$ then $\pi$ is induced by a permutation on $[m]$.
\end{lemma}
\begin{proof}
Consider the set: $T_1 := \{(i, S): i \in \pi(S), S \in \mathcal{H}\}$, which number $|T_1| = \sum_{S \in \mathcal{H}} |\pi(S)| \geq \sum_{S \in \mathcal{H}} |S| = \sum_{i \in [m]} \deg_\mathcal{H}(i) = mr$. Note that \eqref{cond} implies $|T_1| \leq \overline m r$, since otherwise pigeonholing the elements of $T_1$ with respect to their possible first indices $[\overline m]$ would imply $\deg_\mathcal{H}(i) > r$ for some $i \in [m]$. Thus, $\overline m \geq m$.

@@

\begin{proof}[Proof of Lem.~\ref{MainLemma}]
%We will show that the bound \eqref{GapUpperBound} trickles through the intersection semi-lattices of $\{\bm{\mathcal{A}}_S\}_{S \in \mathcal{H}}$ and $\{\bm{\mathcal{B}}_{\overline S}\}_{\overline S \in \pi(\mathcal{H})}$ to yield \eqref{MainLemmaBPD} by virtue of the SIP.  
We begin by showing that $\dim(\bm{\mathcal{B}}_{\pi(S)}) = \dim(\bm{\mathcal{A}}_S)$ for all $S \in \mathcal{H}$. Note that since $\|\mathbf{A}\mathbf{x}\|_2 \leq \max_j\|\mathbf{A}_j\|_2\|\mathbf{x}\|_1$ and $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for all $k$-sparse $\mathbf{x}$, by \eqref{Ldef} we have $L_2(\mathbf{A}) \leq \max_j\|\mathbf{A}_j\|_2$ and therefore (as $\xi > 0$), the right-hand side of \eqref{GapUpperBound} is less than one. It follows from \eqref{dimLem} that $|\pi(S)| \geq \dim(\bm{\mathcal{B}}_{\pi(S)}) \geq \dim(\bm{\mathcal{A}}_S) = |S|$, with the equality by injectivity of $\mathbf{A}_S$. Given that $|S| = |\pi(S)|$, the required fact then follows. Note therefore that $\mathbf{B}_{\pi(S)}$ has full-column rank for all $S \in \mathcal{H}$.

>>>>

\begin{definition}\label{maindef}
Fix $Y = \{ \mathbf{y}_1, \ldots, \mathbf{y}_N\} \subset \mathbb{R}^n$. We say $Y$ has a \textbf{$k$-sparse representation in $\mathbb{R}^m$} if there exists a matrix $\mathbf{A}$ and $k$-sparse $\mathbf{x}_1, \ldots, \mathbf{x}_N \in \mathbb{R}^m$ such that $\mathbf{y}_i = \mathbf{A}\mathbf{x}_i$ for all $i$. 
This representation is \textbf{stable} if for every $\delta_1, \delta_2 \geq 0$, there exists some $\varepsilon = \varepsilon(\delta_1, \delta_2)$ that is strictly positive for positive $\delta_1$ and $\delta_2$ such that if $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^m$ satisfy:
\begin{align*}
	\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \varepsilon(\delta_1, \delta_2),\ \ \   \text{for all $i=1,\ldots,N$},
\end{align*}
then there is some permutation matrix $\mathbf{P}$ and invertible diagonal matrix $\mathbf{D}$ such that for all $i, j$:
\begin{align}\label{def1}
\|\mathbf{A}_j - \mathbf{BPD}_j\|_2 \leq \delta_1 \ \ \text{and} \ \ \|\mathbf{x}_i - \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\overline x}_i\|_1 \leq \delta_2.
\end{align}
\end{definition}

To see how Def. \ref{maindef} directly relates to Prob. \ref{InverseProblem}, suppose that $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ and fix $\delta_1, \delta_2$ to be the desired recovery accuracy in \eqref{def1}. Consider any dataset $Z$ generated as in \eqref{LinearModel} with $\eta \leq \frac{1}{2} \varepsilon(\delta_1, \delta_2)$. Then from the triangle inequality, it follows that any matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^m$ solving Prob.~\ref{InverseProblem} are necessarily within $\delta_1, \delta_2$ of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$, respectively.

@@

\begin{problem}\label{OptimizationProblem}
Find $\mathbf{B}$ and \mbox{$\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^{\overline m}$} that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\overline x}_i\|_0 \ \ 
\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta, \ \text{for all $i$}.
\end{align}
\end{problem}
We then sketch proofs of probabilistic extensions of Thm.~\ref{DeterministicUniquenessTheorem} to random data and dictionaries (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}). Finally, in Sec.~\ref{Discussion}, we discuss both theoretical and practical applications of our main mathematical findings.

@@

Next, we describe a quantitative spark condition. The \emph{lower bound} of a matrix $\mathbf{M} \in \mathbb R^{n \times m}$ is the largest $\alpha$ with \mbox{$\|\mathbf{M}\mathbf{x}\|_2 \geq \alpha\|\mathbf{x}\|_2$} for all $\mathbf{x} \in \mathbb{R}^m$ \cite{Grcar10}. By compactness of the unit sphere, injective linear maps have nonzero lower bound; hence, if $\mathbf{M}$ satisfies \eqref{SparkCondition}, then each submatrix formed from $2k$ of its columns or less has a strictly positive lower bound.

@@

Every $n \times \overline m$ matrix $\mathbf{B}$ for which there are $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N$ satisfying \mbox{$\|\mathbf{A}\mathbf{x}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \varepsilon$} for all $i$ has $\overline m \geq m$ and, provided $(r-1) \overline m < mr$,
\begin{align}\label{Cstable}
\|\mathbf{A}_j- \mathbf{B}_{\overline J} \mathbf{PD}_j\|_2 \leq C_1 \varepsilon, \ \ \text{for all } j \in J,
\end{align}
%
for some nonempty $\overline J \subseteq [\overline m]$ and $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$, permutation matrix $\mathbf{P}$, and and invertible diagonal matrix $\mathbf{D}$.

@@

One can also easily verify that for every $k < m$, there are regular $k$-uniform hypergraphs with the SIP besides the obvious $\mathcal{H} = {[m] \choose k}$; for instance, take $\mathcal{H}$ to be the consecutive intervals of length $k$ in some cyclic order on $[m]$. In this case, a direct consequence of Cor.~\ref{DeterministicUniquenessCorollary} is rigorous verification of the lower bound \mbox{$N = m(k-1){m \choose k} + m$} for sufficient sample size from the introduction. Often, the SIP is achievable with fewer supports. For example, when $k = \sqrt{m}$, take $\mathcal{H}$ to be the $2k$ rows and columns formed by arranging $[m]$ into a square grid.

@@

%\cite{li2004analysis, Georgiev05, Aharon06, Hillar15}
We furthermore note that unlike in previous works, the matrix $\mathbf{A}$ need not satisfy the spark condition to be recoverable from data. As a simple example for $k=2$, let $\mathbf{A} = [ \mathbf{e}_1, \ldots, \mathbf{e}_5, \mathbf{v}]$ where $\mathbf{v} = \mathbf{e}_1 + \mathbf{e}_3 + \mathbf{e}_5$, and take $\mathcal{H}$ to be all consecutive pairs of $[m]$ arranged in cyclic order. Then, dictionary $\mathbf{A}$ satisfies the assumptions of Thm.~\ref{DeterministicUniquenessTheorem} underlying \eqref{Cstable} without satisfying \eqref{SparkCondition}. %(since $\{ \mathbf{A}_1, \mathbf{A}_3, \mathbf{A}_5, \mathbf{A}_6\}$ is not a linearly independent set).
%This weakening allows for a practical (polynomial) amount of data to still guarantee a stable dictionary in the case where the set of sparse supports for the $\mathbf{\overline x}_i$ is known to have a size that grows polynomially in $m$ and $k$.

@@

\begin{proposition}\label{neuroprop}
Sparse neural population activity is recoverable from noisy random linear compression by any method that solves Prob.~\ref{DeterministicUniquenessTheorem} or \ref{SLCopt}; in particular, via biophysically plausible unsupervised sparse linear coding (e.g., \cite{rehnsommer2007, rozell2007neurally, pehlevan2015normative}).
\end{proposition}
%\cite{rehnsommer2007, rozell2007neurally, ganguli2012compressed, hu2014hebbian}

@@

\begin{definition}\label{FriedrichsDefinition}
For a collection of real subspaces $\mathcal{V} = \{V_i\}_{i=1}^\ell$, define $\xi^2 := 0$ when $|\mathcal{V}| = 1$, and otherwise:
\begin{align}\label{xi}
\xi^2(\mathcal{V}) := 1 -  \max \prod_{i=1}^{\ell-1} \sin^2  \theta \left(V_i, \cap_{j>i} V_j \right) ,
\end{align} 
%
where the maximum is taken over all orderings 
%\footnote{We modify the quantity in \cite{Deutsch12} in this way since the subspace ordering is irrelevant to our purpose.} 
of the $V_i$ and the angle $\theta \in (0,\frac{\pi}{2}]$ is defined implicitly as \cite[Def.~9.4]{Deutsch12}:
\begin{align*}
\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \substack{ \mathbf{u} \in U \cap (U \cap W)^\perp, \ \|\mathbf{u}\|_2 \leq 1 \\ \mathbf{w} \in W \cap (U \cap W)^\perp, \  \|\mathbf{w}\|_2 \leq 1 } \right\}.
\end{align*}
\end{definition}
Note that $\theta \in (0,\frac{\pi}{2}]$ implies $0 \leq \xi < 1$, and that $\xi(\mathcal{V}') \leq \xi(\mathcal{V})$ when $\mathcal{V}' \subseteq \mathcal{V}$.\footnote{We acknowledge the counter-intuitive property that $\theta =  \pi/2$ when $U \subseteq W$.}  
The constant $C_2$ in Lem.~\ref{MainLemma} is then:  
\begin{align}\label{Cdef2}
	C_2(\mathbf{A}, \mathcal{H}) := \frac{ r \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ 1- \max_{\mathcal{G} \in {\mathcal{H} \choose r}} \xi( \bm{\mathcal{A}}_\mathcal{G} ) },
\end{align}
% C_2(\mathbf{A}, \mathcal{H}) := \frac{ (r+1) \max_{j \in [m]} \|\mathbf{A}_j\|_2}{ \min_{\mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1}} \xi( \{ \bm{\mathcal{A}}_S \}
which we remark is consistent with the assumption on $C_1$ in the proof of the case $k=1$ at the beginning of this section.

@@

Since there are more than $(k-1){\overline m \choose k}$ vectors $\mathbf{x}_i$ supported in each $S \in \mathcal{H}$, the pigeonhole principle gives $\overline S \in {[\overline m] \choose k}$ and a set of $k$ indices $K \subseteq I(S)$ with all $\mathbf{\overline x}_i$, $i \in K$, supported in $\pi(S) := \overline S$.

@@

\begin{lemma}\label{spanIntersectionLemma}
If $f: U \to V$ is an injective function, then $f\left(\cap_{i=1}^\ell U_i \right) =  \cap_{i=1}^\ell f\left(U_i\right)$ for any $U_1, \ldots, U_\ell \subseteq U$. ($f(\emptyset):=\emptyset$.)
%In particular, if for some $E \in 2^{[m]}$ the map $M \in \mathbb{R}^{n \times m}$ is injective on $\cup_{S \in \mathcal{H}} \text{span}\{e_i\}_{i \in S}$ then $\text{span}\{ M_{\cap_{S \in \mathcal{H}} S} \} = \cap_{S \in \mathcal{H}} \text{span}\{M_S\}$.
\end{lemma}
\begin{proof}
By induction, it is enough to prove the trivial case $\ell = 2$. %Clearly, for any map $f$, if $w \in f(U \cap V)$ then $w \in f(U)$ and $w \in f(V)$, hence $w \in f(U) \cap f(V)$. If $w \in f(U) \cap f(V)$ then $w \in f(U)$ and $w \in f(V)$, hence $w = f(u) = f(v)$ for some $u \in U$ and $v \in V$, implying $u = v$ by injectivity of $f$. Hence $u \in U \cap V$, and $w \in f(U \cap V)$.
\end{proof}
% \vspace{-.4 cm}

\begin{lemma}\label{DistanceToIntersectionLemma}
Let $\mathcal{V} = \{V_i\}_{i=1}^k$ be a set of two or more subspaces of $\mathbb{R}^m$ and let $V = \cap_{i = 1}^k V_i$. For  $\mathbf{x} \in \mathbb{R}^m$, we have (recall Defs.~\ref{dDef}~\&~\ref{FriedrichsDefinition}):
\begin{align}\label{DTILeq}
\text{\rm dist}(\mathbf{x}, V) \leq \frac{1}{1 - \xi(\mathcal{V})} \sum_{i=1}^k \text{\rm dist}(\mathbf{x}, V_i).
\end{align}
\end{lemma}
\vspace{-.3 cm}
\begin{proof} 
% When $V = \{\mathbf{0}\}$, the result is trivial, so suppose otherwise.  
Recall the projection onto the subspace $V \subseteq \mathbb{R}^m$ is the mapping $\Pi_V: \mathbb{R}^m \to V$ that associates with each $\mathbf{x}$ its unique nearest point in $V$; i.e., $\|\mathbf{x} - \Pi_V\mathbf{x}\|_2 = \text{\rm dist}(\mathbf{x}, V)$.
Next, observe:
\begin{align}\label{f}
\|\mathbf{x} - \Pi_V\mathbf{x}\|_2 &\leq \|\mathbf{x} - \Pi_{V_k} \mathbf{x}\|_2 + \|\Pi_{V_k}  \mathbf{x} - \Pi_{V_k}\Pi_{V_{k-1}}\mathbf{x}\|_2 \nonumber \\
&\ \ \ + \cdots + \|\Pi_{V_k} \Pi_{V_{k-1}}\cdots \Pi_{V_1} \mathbf{x} - \Pi_V \mathbf{x}\|_2 \nonumber \\
&\leq \|\Pi_{V_k}\cdots\Pi_{V_{1}} \mathbf{x} - \Pi_V \mathbf{x}\|_2 + \sum_{\ell=1}^k \|\mathbf{x} - \Pi_{V_{\ell}} \mathbf{x}\|_2,
\end{align}
using the triangle inequality and that the spectral norm satisfies $\|\Pi_{V_{\ell}}\|_2 \leq 1$ for all $\ell$ (since $\Pi_{V_{\ell}}$ are orthogonal projections).

@@

Finally, substituting this into \eqref{f} and rearranging produces \eqref{DTILeq} after substituting $\xi(\mathcal{V})$ for $z$.
\end{proof}
\begin{lemma}\label{NonEmptyLemma} 
Fix an $r$-regular hypergraph $\mathcal{H} \subseteq 2^{[m]}$ satisfying the SIP. If the map $\pi: \mathcal{H} \to 2^{[\overline m]}$ has $\sum_{S \in \mathcal{H}} |\pi(S)| \geq \sum_{S \in \mathcal{H}} |S|$ and:
\begin{align}\label{cond}
	|\cap \pi(\mathcal{G})| \leq |\cap \mathcal{G} |,\ \ \   \text{for } \mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1},
\end{align}
%
then $\overline m \geq m$, and if $(r-1) \overline m < mr$ then the map $i \mapsto \cap \pi(\sigma(i))$ is an injective map to $[\overline m]$ from some $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$. %In particular, if $\overline m = m$ then $\pi$ is induced by a permutation on $[m]$.
\end{lemma}
\begin{proof}
Consider the set: $T_1 := \{(i, S): i \in \pi(S), S \in \mathcal{H}\}$, which number $|T_1| = \sum_{S \in \mathcal{H}} |\pi(S)| \geq \sum_{S \in \mathcal{H}} |S| = \sum_{i \in [m]} \deg_\mathcal{H}(i) = mr$. Note that \eqref{cond} implies $|T_1| \leq \overline m r$, since otherwise pigeonholing the elements of $T_1$ with respect to their possible first indices $[\overline m]$ would imply $\deg_\mathcal{H}(i) > r$ for some $i \in [m]$. Thus, $\overline m \geq m$.

@@

\begin{proof}[Proof of Lem.~\ref{MainLemma}]
%We will show that the bound \eqref{GapUpperBound} trickles through the intersection semi-lattices of $\{\bm{\mathcal{A}}_S\}_{S \in \mathcal{H}}$ and $\{\bm{\mathcal{B}}_{\overline S}\}_{\overline S \in \pi(\mathcal{H})}$ to yield \eqref{MainLemmaBPD} by virtue of the SIP.  
We begin by showing that $\dim(\bm{\mathcal{B}}_{\pi(S)}) = \dim(\bm{\mathcal{A}}_S)$ for all $S \in \mathcal{H}$. Note that since $\|\mathbf{A}\mathbf{x}\|_2 \leq \max_j\|\mathbf{A}_j\|_2\|\mathbf{x}\|_1$ and $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for all $k$-sparse $\mathbf{x}$, by \eqref{Ldef} we have $L_2(\mathbf{A}) \leq \max_j\|\mathbf{A}_j\|_2$ and therefore (as $\xi > 0$), the right-hand side of \eqref{GapUpperBound} is less than one. From \eqref{dimLem}, we have $|\pi(S)| \geq \dim(\bm{\mathcal{B}}_{\pi(S)}) \geq \dim(\bm{\mathcal{A}}_S) = |S|$, with the equality by injectivity of $\mathbf{A}_S$. As $|\pi(S)| = |S|$, the claim follows. Note, therefore, that $\mathbf{B}_{\pi(S)}$ has full-column rank for all $S \in \mathcal{H}$.
